Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-12 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Continue to fill gaps in Ch 4 and 6

Figures~\ref{IWGB_HF_Comp_Center} and \ref{IWGB_HF_Comp_Edge} present the averaged peak values from the two heat flux gauge arrays.  For the gasoline fire experiment at the 0.08~m position the wall temperatures and the heat flux values from the burner edge array had values that were higher than when those measured at the 0.0~m position. The peak heat release rates for the positions at 0.0~m, 0.08~m and 0.15~m were 120~kW, 110~kW, and 110~kW respectively.  So it does not seem that a change in HRR was the source of the increase.

@@

@@

In the radiant flooring panel experiments the gypsum wallboard was exposed to an incident heat flux that ranged from approximately 21~kW/m$^2$ to 2~kW/m$^2$ over the length of the sample. In addition to the heat flux from the radiant panel, there is a small flame front approximately 10 to 30~mm in height that serves as the pilot along the length of the sample and provides additional energy to the fuel immediately ahead of the flame.  The impact is limited to a small distance ahead of the flame, estimated to be less than 5~mm, since the flame is drawn back to the leading edge of the sample (burned portion) by the draft of the radiant panel.  This effect is shown in Figure~\ref{FRP_flamefront}. In Figure~\ref{FRP_flamefront} several thermocouples can be seen. The exposed-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm nominal diameter, were installed by pulling the positive and negative wires through holes drilled through the gypsum wallboard sample and positioning the bead of the thermocouple in a slit in the painted paper face made with a razor blade.  This was the same installation method that was used in the instrumented gypsum board wall experiments and in the cone calorimeter samples to measure ignition temperature.  The thermocouples were installed in the center of the sample at 100~mm intervals beginning at 100~mm from the edge of the sample closest to the radiant panel.  Three replicate experiments were conducted.  The average time that the flame front reached the thermocouples and the temperature of the thermocouple just prior to the painted paper surface of the gypsum wallboard sample igniting is presented in Table~\ref{tab:FRP_timetemp}.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/FRP_flamefront} \\
	\caption[Photograph of the flame front burning on the gypsum wallboard sample in the flooring radiant panel] {Photograph of the flame front burning on the gypsum wallboard sample in the flooring radiant panel. Thermocouples imbedded at 100~mm intervals in the surface of the painted paper can be seen in front of and behind the flame front. The flame is leaning back in the direction the radiant panel.}
	\label{FRP_flamefront}
\end{figure}

 With the flooring radiant panel test, the incident heat flux from the panel and the pre-heated gypsum wallboard sample enables flame spread for approximately 500~s.  The temperatures prior to ignition ranged from an average of 315 $^{\circ}$C for the 100~mm position to 242 $^{\circ}$C for the 400~mm position.  The average

>>>>

Figures~\ref{IWGB_HF_Comp_Center} and \ref{IWGB_HF_Comp_Edge} present the averaged peak values from the two heat flux gauge arrays.  For the gasoline fire experiment at the 0.08~m position the wall temperatures and the heat flux values from the burner edge array had values that were higher than when those measured at the 0.0~m position. The peak heat release rates for the positions at 0.0~m, 0.08~m and 0.15~m were within the average peak heat release rate range of 80~kW~$\pm$~24~kW.  So a change in HRR was not the source of the increase.  The increased temperatures and heat flux at this position appears to be a matter of the flame movement toward the wall.  In general, once the burner position for either natural gas or gasoline was within 0.15~m, or half the distance of the burner side dimension, the flames tended to pull toward the wall. This motion was not a steady motion but an oscillatory one.

@@

Figure~\ref{IWGBGAS6_seq} shows a comparison of a series eight photographs from the experiment where the gasoline burner was positioned 0.15~m from the instrumented gypsum wallboard.  The photographs capture both the oscillatory motion of the flame toward the wall and the burning of the gases which are pyrolyzing off of the wall and result in the formation of the lower edge of the fire pattern.

\begin{figure}[p]
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq1}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq2}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq3}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq4} \\

	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq5}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq6}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq7}
	\includegraphics[width=1.3in]{../Figures/IWGBGAS6_Seq8} \\

	\caption[Photographs of the fire pattern formation from the gasoline burner positioned 0.15~m from the wall]{Photographic sequence of the fire pattern formation from the gasoline burner positioned 0.15~m from the instrumented gypsum wallboard.}
	\label{IWGBGAS6_seq}
\end{figure}

@@

In the radiant flooring panel experiments the gypsum wallboard was exposed to an incident heat flux that ranged from approximately 21~kW/m$^2$ to 2~kW/m$^2$ over the length of the sample. In addition to the heat flux from the radiant panel, there is a small flame front approximately 10 to 30~mm in height that serves as the pilot along the length of the sample and provides additional energy to the fuel immediately ahead of the flame.  The impact is limited to a small distance ahead of the flame, estimated to be less than 5~mm, since the flame is drawn back to the leading edge of the sample (burned portion) by the draft of the radiant panel.  This effect is shown in Figure~\ref{FRP_flamefront_arrows}. In Figure~\ref{FRP_flamefront_arrows} several thermocouples can be seen. The exposed-bead, Chromel-Alumel (type K) thermocouples, with a 1.0~mm nominal diameter, were installed by pulling the positive and negative wires through holes drilled through the gypsum wallboard sample and positioning the bead of the thermocouple in a slit in the painted paper face made with a razor blade.  This was the same installation method that was used in the instrumented gypsum board wall experiments and in the cone calorimeter samples to measure ignition temperature.  The thermocouples were installed in the center of the sample at 100~mm intervals beginning at 100~mm from the edge of the sample closest to the radiant panel.  Three replicate experiments were conducted.  The average time that the flame front reached the thermocouples and the temperature of the thermocouple just prior to the painted paper surface of the gypsum wallboard sample igniting is presented in Table~\ref{tab:FRP_timetemp}.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/FRP_flamefront_arrows} \\
	\caption[Photograph of the flame front burning on the gypsum wallboard sample in the flooring radiant panel] {Photograph of the flame front burning on the gypsum wallboard sample in the flooring radiant panel. Thermocouples imbedded at 100~mm intervals in the surface of the painted paper can be seen in front of and behind the flame front. The flame is leaning back in the direction the radiant panel.}
	\label{FRP_flamefront_arrows}
\end{figure}

 With the flooring radiant panel test, the incident heat flux from the panel and the pre-heated gypsum wallboard sample enables flame spread for approximately 500~s.  The temperatures prior to ignition ranged from an average of 315 $^{\circ}$C for the 100~mm position to 242 $^{\circ}$C for the 400~mm position.
